---
title: "DAB_303 Marketing Analytics"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r}
library(tidyverse)
```


## Implementation of Customer Segmentation

Reading in the dataset "customer_data".
```{r}
customer_data <- read.csv(file.choose())

str(customer_data)
```

Viewing the column names included by the dataset.
```{r}
names(customer_data)
```

View the first 6 rows of the dataset using head() function.
```{r}
head(customer_data)
```

Summarizing the columns "Age" and "Annual.Income..k.." and find standard deviations.
```{r}
#Summarizing the "Age" column
summary(customer_data$Age)

#Standard deviation of Age
sd(customer_data$Age)

#Summarizing the column "Annual.Income..k.."
summary(customer_data$Annual.Income..k..)

#Standard deviation pf column "Annual.Income..k.."
sd(customer_data$Annual.Income..k..)

```

Find standard deviation of "Spending.Score..1.100." column.
```{r}
sd(customer_data$Spending.Score..1.100.)
```


## Customer Gender Visualization

Plot a barplot and a piechart to show the gender distribution across our "customer_data" dataset

```{r}
a = table(customer_data$Gender)
barplot(a, main = "Using BarPlot to display Gender Comparision",
           ylab = "Count",
           xlab = "Gender",
           col = rainbow(2),
           legend=rownames(a) )
```

```{r}
pct = round(a/sum(a)*100)
lbs = paste(c("Female","Male")," ",pct,"%",sep=" ")
library(plotrix)
pie3D(a, labels=lbs, 
         main = "Pie Chart Depicting Ratio of Female and Male")
```

## Visualization of Age Distribution

Plotting a histogram to view the distribution to plot the frequency of customer ages.
```{r}
hist(customer_data$Age,
     col = "Magenta",
     main = "Histogram to Show Count of Age Class",
     xlab = "Age Class",
     ylab = "Frequency",
     labels = TRUE)
```

Plot boxplot for descriptive analysis of "Age" column.
```{r}
boxplot(customer_data$Age,
        col = "orange",
        main = "Boxplot for Descriptive Analysis of Age")
```

## Analysis of the Annual Income of the Customers

Get the summary and then plotting a histogram to analyze the annual income.
```{r}
summary(customer_data$Annual.Income..k..)

hist(customer_data$Annual.Income..k..,
  col="#660033",
  main="Histogram for Annual Income",
  xlab="Annual Income Class",
  ylab="Frequency",
  labels=TRUE)
```

We will proceed to examine this data using a density plot.
```{r}
plot(density(customer_data$Annual.Income..k..),
    col="blue",
    main="Density Plot for Annual Income",
    xlab="Annual Income Class",
    ylab="Density")
polygon(density(customer_data$Annual.Income..k..),
        col="yellow")
```

## Analyzing Spending Score of the Customers

Creating summary and boxplot for anlyzing spending score of customers.
```{r}
summary(customer_data$Spending.Score..1.100.)

boxplot(customer_data$Spending.Score..1.100.,
   horizontal=TRUE,
   col="#990000",
   main="BoxPlot for Descriptive Analysis of Spending Score")
```

Now, plotting a histogram for the spending score data of customers.
```{r}
hist(customer_data$Spending.Score..1.100.,
    main="HistoGram for Spending Score",
    xlab="Spending Score Class",
    ylab="Frequency",
    col="#6600cc",
    labels=TRUE)
```


## Determining Optimal Clusters

### Elbow Method

we proceed to plot iss(sum of square) based on the number of k clusters. This plot denotes the appropriate number of clusters required in our model. In the plot, the location of a bend or a knee is the indication of the optimum number of clusters.
```{r}
library(purrr)

set.seed(123)

# function to calculate total intra-cluster sum of square 
iss <- function(k) {
  kmeans(customer_data[,3:5],k,iter.max=100,nstart=100,algorithm="Lloyd" )$tot.withinss
}
k.values <- 1:10
iss_values <- map_dbl(k.values, iss)
plot(k.values, iss_values,
    type="b", pch = 19, frame = FALSE, 
    xlab="Number of clusters K",
    ylab="Total intra-clusters sum of squares")
```

## Gap Statistic Method

We can utilize the clusGap function for providing gap statistic as well as standard error for a given output.
```{r}
library(cluster)          #clustering algorithms
library(factoextra)       #clustering algorithms & visualization

set.seed(125)
stat_gap <- clusGap(customer_data[,3:5], FUN = kmeans, nstart = 25,
            K.max = 10, B = 50)
fviz_gap_stat(stat_gap)
```


### Now, let us take k = 6 as our optimal cluster –
```{r}
k6<-kmeans(customer_data[,3:5], 6, iter.max=100, nstart=50, algorithm="Lloyd")
k6
```


## Visualizing the Clustering Results using the First Two Principle Components

```{r}
pcclust=prcomp(customer_data[,3:5],scale=FALSE)              #principal component analysis

summary(pcclust)
pcclust$rotation[,1:2]
```

```{r}
set.seed(1)
ggplot(customer_data, aes(x =Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
              breaks=c("1", "2", "3", "4", "5","6"),
              labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
```

```{r}
ggplot(customer_data, aes(x =Spending.Score..1.100., y =Age)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                      breaks=c("1", "2", "3", "4", "5","6"),
                      labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
```

```{r}
kCols=function(vec){cols=rainbow (length (unique (vec)))
return (cols[as.numeric(as.factor(vec))])}
digCluster<-k6$cluster; dignm<-as.character(digCluster);         # K-means clusters

plot(pcclust$x[,1:2], col =kCols(digCluster),pch =19,xlab ="K-means",ylab="classes")
legend("bottomleft",unique(dignm),fill=unique(kCols(digCluster)))
```




